\bf DOGE's Downsizing, Can AI Read the Reddit Room?

### 2025-04-09 code/ml_llm/R/assign_3.qmd
2nd round of assing 3 paper

---
title: "DOGE's Downsizing, Can AI Read the Reddit Room?"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default
date: last-modified
author:
  - name: Kevin Linares
    affiliations:
      - name: University of Maryland
  - name: Felix Baez-Santiago
  - name: Aria Lu
  - name: Gloria Zhou
    affiliations:
      - name: University of Michigan
abstract: |
 We investigate public sentiment on Reddit regarding the Department of Government Efficiency's federal workforce reduction by classifying 400 labeled comments (28% favor, 18% neutral, 54% oppose) using supervised and unsupervised Large Language Models. Supervised models showed moderate success, particularly with "oppose" comments, but struggled with "favor" and "neutral" stances. Similarly, LLMs best identified "oppose" sentiment but exhibited low precision for "favor" and "neutral." These findings highlight the challenges of accurately gauging nuanced public opinion on government policy changes using automated methods on social media data.

@@

# read in ML results from repo
results_folder <- "~/repos/SURV622_Assignment/outputs/ml_llm/"

@@

## Methods {#sec-meth}

*Data*.
We collected Reddit comments from subreddits related to the topic of interest in early March of 2024.
This resulted in 12,553 comments which we preprocessed by removing web-URLs, replace special characters (i.e., replaced "\@" with "at"), replaced numeric values with their spelling, and confirmed the absence of duplicate comments.
We then randomly selected 400 comments (without replacement) and assigned them to four graduate students for coding, categorizing each comment as favoring, opposing, or neutral towards DOGE's approach to federal workforce reductions.
Table 1 presents the breakdown in percent of our the comments we labelled and later use to train and evaluate our machine learning models.
A review of the comments revealed more negative opinionated statements and reactions.

@@

We built two supervised machine learning models–K-nearest neighbor and random forest–to classify Reddit comments concerning stance on the current reduction in federal workforce into favor, oppose, or neutral.
For predictors we used the Reddit score, up-votes, and down-votes each comment received at the time of the collection.
Additionally, we used the keyword that was used to scrape these data along with the subreddit where the comment was posted.
**DISCUSS ADDITIONAL FEATURE ENGINERRING SUCH AS TEXT EDITING THAT WENT INTO THE MODEL.** For the KNN model, we set the number of neighbors hyperparameter to three, and for the random forest, mtry was kept at a constant two as we did not have many features in these models.
We took an 80/20 split for our training-testing sets that we used to train each model and evaluate.

In addition to our supervised machine learning models, we also applied to our comments two LLMs, gemma 3.12 and llama 3.2.
3B.
Gemma, was developed by Google and is based on Gemini, is able to generate text output using instruction-tuned prompts from the user, such as providing answers on a user specified task reviewing text.
We deployed gemma on a higher performance computing cluster ([Great Lakes HCP](https://its.umich.edu/advanced-research-computing/)) and processed text from Reddit comments on a GPU.
We were also interested in deploying other lightweight LLMs locally on a machine with a dedicated-GPU and found llama, developed by Meta AI, to be an excellent candidate.
Llama excels at text classification which is how we used it to classify the stance for each comment text given a user specified prompt.
We choose both of these LLMs as candidates to test against our supervised approaches because of their ability to apply locally and reputable performance on zero-shot classification tasks.
Both of these models received the same prompt and tasked which were developed by the research team to reflect the utility of classifying stance within our topic of interest.

|  |  |
|----------------|--------------------------------------------------------|
| **Prompt**: | "Is this comment in 'favor', 'neutral', or 'oppose' the reduction in federal workforce? Provide one word answer only! |
| **Task**: | "You have assumed the role of a stakeholder that is presented with a reddit comment from a likely federal worker related to the current policies on reducing the federal workforce. Please determine the author's stance on this topic, and only provide the answer." |

We evaluate model performance using the F1-score which is a harmonic mean balance between precision and recall.
This metric considers both the ability to correctly identify positive instances, known as recall, and the ability to avoid incorrectly labeling negative instances as positive (i.e., precision).
A high F1-score indicates a model with well balanced recall and precision that both minimize false negativesand false positives.
The second performance metric we used is recall, as this determines how many of the actual positive cases the model can correctly identify.
Recall can be prone to incorrectly labeling negative cases as positive, suggesting more false positives, which is why we intereggoate on two performance metrics in this study.

>>>>

---
title: "DOGE's Downsizing, Can AI Read the Reddit Room?"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default
date: last-modified
author:
  - name: Kevin Linares
    affiliations:
      - name: University of Maryland
  - name: Felix Baez-Santiago
  - name: Aria Lu
  - name: Gloria Zhou
    affiliations:
      - name: University of Michigan
abstract: |
  We investigate public sentiment on Reddit regarding the Department of Government Efficiency's federal workforce reduction by classifying 400 labeled comments (28% favor, 18% neutral, 54% oppose) using supervised and unsupervised Large Language Models (LLM). Supervised models showed moderate success, particularly with "oppose" comments, but struggled with "favor" and "neutral" stances. Similarly, LLMs best identified "oppose" sentiment but exhibited low precision for "favor" and "neutral." These findings highlight the challenges of accurately gauging nuanced public opinion on government policy changes using automated methods on social media data.

@@

# read in ML results from repo
results_folder <- "~/work/d/SURV622_Assignment/outputs/ml_llm/"

@@

## Method {#sec-meth}

We collected a total of 12,553 Reddit comments in early March 2024 from subreddits relevant to the topic of interest.
These comments were retrieved using keyword-based queries designed to capture posts of discussions that would reflect users’ stances on the federal workforce reduction by DOGE.
To prepare the data for analysis, a comprehensive preprocessing pipline were conducted in Pytho 3.
This process involved removing HTML tags and web-URLs, stripping special characters and tagging artifacts such as “\@”, and replacing emojis with textual descriptions.
Also, it entailed eliminating punctuations and converting numerical values into their spelled-out forms.
The text was standardized to lowercase, lemmatized using the SpaCy library, and further refined by removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while Duplicate comments were identified and removed to maintain data integrity.

From this cleaned dataset, 400 uniqe comments without replacement were randomly sampled for manual annotation.
Four graudate students independently coded and labeled each comment with one of three possible stances:favor, opopose, or neutral toward DOGE’s approach to federal workforce reductions.
As shown in Table 1, there were 28% favor, 18% neutral, and 54% oppose.
These labeled comments were subsequently used for training and evaluating both supervised and unsupervised classification models.

@@

The feature set used in model training and evluation consisted of two primary categories: metadata features and textual features.
Metadata features included the Reddit comment’s score, the subreddit in which it appeared in, and the search term that retrieved it.
As the variables for score and upvotes were perfectly collinear in this dataset, only the score was retained.
To ensure all values were non-negative, the score was adjusted by subtracting the dataset’s minimum value from each score.
The categorical variables, subreddit and search term, were numerically encoded using scikit-learn’s label encoder.

Textual features were derived from each comment and represented using two distinct approaches: unigram term frequency and term freuqency-inverse document frequency (TF-IDF).
The unigram used a bag-of-words model to represent each word as a feature based on its frequency in comment.
The TF-IDF transformation built upon these raw counts by accounting for term importance across the corpus.
For both unigram and TF-IDF methods, the total number of text-derived features was 2,102.
Combined with the three metadata features, each dataset contained a total of 2,105 features.
Two datasets were constructed for classification: one combining the metadata with unigram features, and another with metadat and TF-IDF features.
Each dataset was divded into 80% training data and 20% test data.
All supervised learning models were implemented using Python 3, with scikit-learn and XGBoost as the primary machine learning (ML) techniques.
Classification was conducted using a diverse set of algorithms, including Extra Trees Classifier, Random Forest Classifier, Logistic Regression, Ridge Classifier, K-Nearest Neightbors (KNN), Nearest Centroid, Multi-layer Perceptron Classifier, Bernoulli Naive Bayes, Multino- mial Naive Bayes, Linear Support Vector Classifier, Decision Tree Classifier, Extra Tree Classifier, and the XGBoost Classifier.

All models were trained using default hyperparameter confiurations.
The number of neighbors for KNN model was set to three, and the mtry parameter for Random Forrest was fixed at two, reflecting the relatively small number of metadata features in those models.
All experiments were conducted on a personal computer without the use of a GPU

In addition to traditional supervised models, we employed two instruction-tuned LLMs to perform zero- shot classification: Gemma 3.12B, developed by Google and based on the Gemini framework, and LLaMA 3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes High-Performance Computing (HPC) cluster, while LLaMA was deployed locally on a GPU-enabled personal workstation using the rollama r pacakge.
These models were selected for their strong zero-shot classification capabilities and ability to be run locally or on cluster environments without reliance on commercial APIs.

|  |  |
|------------------|------------------------------------------------------|
| **Prompt**: | "Is this comment in 'favor', 'neutral', or 'oppose' the reduction in federal workforce? Provide one word answer only! |
| **Task**: | "You have assumed the role of a stakeholder that is presented with a reddit comment from a likely federal worker related to the current policies on reducing the federal workforce. Please determine the author's stance on this topic, and only provide the answer." |

The model performance was evaluated by the F1-score and recall.
The F1-score, a harmonic mean of precision and recall, serves as a balanced indicator of a model’s ability to correctly identify relevant instances while minimizing both false positives and false negatives.
Recall was also independently assessed to evaluate the proportion of true postive cases accurately identified by each model.
While a high recall may increase the risk of false positives, its combination with the F1-score allows a more robust evaluation of model performance, especially for addressing the challenges inherent in imbalanced classification tasks

### 2025-04-09 code/ml_llm/R/assign_3.tex
2nd round of assing 3 paper

\date{April 8, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigate public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models. Supervised models
showed moderate success, particularly with ``oppose'' comments, but
struggled with ``favor'' and ``neutral'' stances. Similarly, LLMs best
identified ``oppose'' sentiment but exhibited low precision for
``favor'' and ``neutral.'' These findings highlight the challenges of
accurately gauging nuanced public opinion on government policy changes
using automated methods on social media data.
\end{abstract}

@@

\section{Methods}\label{sec-meth}

\emph{Data}. We collected Reddit comments from subreddits related to the
topic of interest in early March of 2024. This resulted in 12,553
comments which we preprocessed by removing web-URLs, replace special
characters (i.e., replaced ``@'' with ``at''), replaced numeric values
with their spelling, and confirmed the absence of duplicate comments. We
then randomly selected 400 comments (without replacement) and assigned
them to four graduate students for coding, categorizing each comment as
favoring, opposing, or neutral towards DOGE's approach to federal
workforce reductions. Table 1 presents the breakdown in percent of our
the comments we labelled and later use to train and evaluate our machine
learning models. A review of the comments revealed more negative
opinionated statements and reactions.

@@

We built two supervised machine learning models--K-nearest neighbor and
random forest--to classify Reddit comments concerning stance on the
current reduction in federal workforce into favor, oppose, or neutral.
For predictors we used the Reddit score, up-votes, and down-votes each
comment received at the time of the collection. Additionally, we used
the keyword that was used to scrape these data along with the subreddit
where the comment was posted. \textbf{DISCUSS ADDITIONAL FEATURE
ENGINERRING SUCH AS TEXT EDITING THAT WENT INTO THE MODEL.} For the KNN
model, we set the number of neighbors hyperparameter to three, and for
the random forest, mtry was kept at a constant two as we did not have
many features in these models. We took an 80/20 split for our
training-testing sets that we used to train each model and evaluate.

In addition to our supervised machine learning models, we also applied
to our comments two LLMs, gemma 3.12 and llama 3.2. 3B. Gemma, was
developed by Google and is based on Gemini, is able to generate text
output using instruction-tuned prompts from the user, such as providing
answers on a user specified task reviewing text. We deployed gemma on a
higher performance computing cluster
(\href{https://its.umich.edu/advanced-research-computing/}{Great Lakes
HCP}) and processed text from Reddit comments on a GPU. We were also
interested in deploying other lightweight LLMs locally on a machine with
a dedicated-GPU and found llama, developed by Meta AI, to be an
excellent candidate. Llama excels at text classification which is how we
used it to classify the stance for each comment text given a user
specified prompt. We choose both of these LLMs as candidates to test
against our supervised approaches because of their ability to apply
locally and reputable performance on zero-shot classification tasks.
Both of these models received the same prompt and tasked which were
developed by the research team to reflect the utility of classifying
stance within our topic of interest.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2222}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7778}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and only provide the answer.'' \\
\end{longtable}

We evaluate model performance using the F1-score which is a harmonic
mean balance between precision and recall. This metric considers both
the ability to correctly identify positive instances, known as recall,
and the ability to avoid incorrectly labeling negative instances as
positive (i.e., precision). A high F1-score indicates a model with well
balanced recall and precision that both minimize false negativesand
false positives. The second performance metric we used is recall, as
this determines how many of the actual positive cases the model can
correctly identify. Recall can be prone to incorrectly labeling negative
cases as positive, suggesting more false positives, which is why we
intereggoate on two performance metrics in this study.

>>>>

\date{April 9, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigate public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models (LLM). Supervised
models showed moderate success, particularly with ``oppose'' comments,
but struggled with ``favor'' and ``neutral'' stances. Similarly, LLMs
best identified ``oppose'' sentiment but exhibited low precision for
``favor'' and ``neutral.'' These findings highlight the challenges of
accurately gauging nuanced public opinion on government policy changes
using automated methods on social media data.
\end{abstract}

@@

\section{Method}\label{sec-meth}

We collected a total of 12,553 Reddit comments in early March 2024 from
subreddits relevant to the topic of interest. These comments were
retrieved using keyword-based queries designed to capture posts of
discussions that would reflect users' stances on the federal workforce
reduction by DOGE. To prepare the data for analysis, a comprehensive
preprocessing pipline were conducted in Pytho 3. This process involved
removing HTML tags and web-URLs, stripping special characters and
tagging artifacts such as ``@'', and replacing emojis with textual
descriptions. Also, it entailed eliminating punctuations and converting
numerical values into their spelled-out forms. The text was standardized
to lowercase, lemmatized using the SpaCy library, and further refined by
removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while
Duplicate comments were identified and removed to maintain data
integrity.

From this cleaned dataset, 400 uniqe comments without replacement were
randomly sampled for manual annotation. Four graudate students
independently coded and labeled each comment with one of three possible
stances:favor, opopose, or neutral toward DOGE's approach to federal
workforce reductions. As shown in Table 1, there were 28\% favor, 18\%
neutral, and 54\% oppose. These labeled comments were subsequently used
for training and evaluating both supervised and unsupervised
classification models.

@@

The feature set used in model training and evluation consisted of two
primary categories: metadata features and textual features. Metadata
features included the Reddit comment's score, the subreddit in which it
appeared in, and the search term that retrieved it. As the variables for
score and upvotes were perfectly collinear in this dataset, only the
score was retained. To ensure all values were non-negative, the score
was adjusted by subtracting the dataset's minimum value from each score.
The categorical variables, subreddit and search term, were numerically
encoded using scikit-learn's label encoder.

Textual features were derived from each comment and represented using
two distinct approaches: unigram term frequency and term
freuqency-inverse document frequency (TF-IDF). The unigram used a
bag-of-words model to represent each word as a feature based on its
frequency in comment. The TF-IDF transformation built upon these raw
counts by accounting for term importance across the corpus. For both
unigram and TF-IDF methods, the total number of text-derived features
was 2,102. Combined with the three metadata features, each dataset
contained a total of 2,105 features. Two datasets were constructed for
classification: one combining the metadata with unigram features, and
another with metadat and TF-IDF features. Each dataset was divded into
80\% training data and 20\% test data. All supervised learning models
were implemented using Python 3, with scikit-learn and XGBoost as the
primary machine learning (ML) techniques. Classification was conducted
using a diverse set of algorithms, including Extra Trees Classifier,
Random Forest Classifier, Logistic Regression, Ridge Classifier,
K-Nearest Neightbors (KNN), Nearest Centroid, Multi-layer Perceptron
Classifier, Bernoulli Naive Bayes, Multino- mial Naive Bayes, Linear
Support Vector Classifier, Decision Tree Classifier, Extra Tree
Classifier, and the XGBoost Classifier.

All models were trained using default hyperparameter confiurations. The
number of neighbors for KNN model was set to three, and the mtry
parameter for Random Forrest was fixed at two, reflecting the relatively
small number of metadata features in those models. All experiments were
conducted on a personal computer without the use of a GPU

In addition to traditional supervised models, we employed two
instruction-tuned LLMs to perform zero- shot classification: Gemma
3.12B, developed by Google and based on the Gemini framework, and LLaMA
3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes
High-Performance Computing (HPC) cluster, while LLaMA was deployed
locally on a GPU-enabled personal workstation using the rollama r
pacakge. These models were selected for their strong zero-shot
classification capabilities and ability to be run locally or on cluster
environments without reliance on commercial APIs.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and only provide the answer.'' \\
\end{longtable}

The model performance was evaluated by the F1-score and recall. The
F1-score, a harmonic mean of precision and recall, serves as a balanced
indicator of a model's ability to correctly identify relevant instances
while minimizing both false positives and false negatives. Recall was
also independently assessed to evaluate the proportion of true postive
cases accurately identified by each model. While a high recall may
increase the risk of false positives, its combination with the F1-score
allows a more robust evaluation of model performance, especially for
addressing the challenges inherent in imbalanced classification tasks
